---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BezELS

<!-- badges: start -->
<!-- badges: end -->

BezELS (Bezier splines for Estimating Limit Sets) performs tail inference for bivariate data with unit exponential marginal distributions. The package implements the methodology from [this paper on arXiv](https://arxiv.org/abs/2306.13257).

## Installation

You can install the development version of BezELS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("reetamm/BezELS")
```

## Example

The following example generates data from a logistic copula, fits a Bezier spline to the limit set boundary using MCMC, and does tail inference.

```{r estimation}
library(BezELS)
set.seed(1)
# Generate some data
simdata             <- gen_data_exp(n = 500, theta = 0.3, tau = 0.75, copula = 'l')
x                   <- simdata$x
r                   <- simdata$r
w                   <- simdata$w
data_marg_r_0       <- simdata$data_marg_r_0
# Run MCMC to fit a bezier spline to the limit set
samples  <- fit_mcmc_bezier( N = data_marg_r_0$N,
                         r = data_marg_r_0$r,
                         w = data_marg_r_0$w,
                         r_0 = data_marg_r_0$r_0,
                         iters = 1100, burn = 100,
                         traceplot=F, print.every = 100)
samples <- samples[101:1100,] # drop the burn-in samples
```
We're going to plot the posterior draws in both polar and Euclidean coordinates.
```{r plots, warning=FALSE}
r_0_marg            <- simdata$r_0_marg
above_thresh_marg   <- simdata$above_thresh_marg
par(mfrow=c(1,2))
medcurve = plot_bezier_polar(mcmc_samples = samples, x = x, r_0_marg = r_0_marg, above_thresh_marg = above_thresh_marg,
                                 copula = 'l', theta = 0.3, thin = 1, plot.fig = T, plottitle = 'Polar coordinates')
plot_bezier_euclidean(mcmc_samples = samples, x = x, theta = 0.3, copula = 'l', plottitle = 'Euclidean coordinates',
                          thin = 1, above_thresh_marg = above_thresh_marg, plot.truth = T, medcurve = medcurve)
```

Finally, let's calculate some metrics.
```{r inference}
summary(apply(samples[,1:10], 1, eta_empirical))
summary(samples[,11])

omega <- seq(0.01,0.99,by = 0.01)
delta <- omega
lambda_omega    <- rep(NA,99)
tau1_delta      <- rep(NA,99)

for(i in 1:99){
    lambda_omega[i] <- median(apply(samples,1,lambda_empirical, omega = omega[i]))
    tau1_delta[i]   <- median(apply(samples,1,tau_empirical, delta = delta[i]))
}
par(mfrow=c(1,2))
plot(omega,lambda_omega,'l', ylab = expression(lambda(omega)), xlab = expression(omega))
plot(delta,tau1_delta,'l', ylab = expression(tau[1](delta)), xlab = expression(delta))
```
